### 2012-01-31 GUIErrorDetector/paper/technique.tex
significant revise

We first give our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error algorithm to different GUI frameworks.

\subsection{Approach Formulation}

This section formulates our high-level approach. We define the
notations of \textit{UI thread}, \textit{non-UI threads},
\textit{UI-accessing methods},
 \textit{safe methods}, and \textit{invalid thread access error}, and
two assumptions we make with regard to error detection.

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specifically, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretically ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph used in practice is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
In an object-oriented programming language like
Java, a method is the basic logic unit in a program. Reporting
potential errors at the method level can be sufficient in most cases.
Thus, we believe using the call graph is well suited for this problem while
compared with other program representations like control flow graph or
program dependence graph.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes for it (line 3). The algorithm starts to traverse
the call graph from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} calls (line 7). Each \CodeIn{Thread.start()} call
indicates that a new, non-UI thread is spawned.  After that, 
the algorithm uses a BFS to find reachable UI-accessing methods (lines 9 -- 23)
from each \CodeIn{Thread.start()} node.
Specifically, it uses a queue to keep track of all visited nodes, and
then checks every successive node to be visited (line 12);
it reports an error if that node
is a UI-accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing otherwise (line 19).

If the algorithm determines that a UI-accessing method may be invoked
by a non-UI thread, it creates an error report containing the full
method call chain from the entry node (line 15, the \CodeIn{createErrorReport}
method). An example call chain for
the motivating example in Figure~\ref{fig:androiderror} is shown
in Figure~\ref{fig:report}.

The algorithm in Figure~\ref{fig:detectalgorithm} uses BFS to search
for potential error-revealing paths. Other graph search strategies
such as Depath-First Search (DFS) or exhaustive path search can also be
adapted. In our evaluation, we empirically show that BFS
is more effective than DFS and exhaustive search in finding errors.

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph 
%for partial programs such as an Eclipse plugin, or
in the presence of reflection is non-trivial. 
In particular, GUI applications built on some popular frameworks like Android
often use configuration files and reflection to specify GUI layout.
%or extension points.
%Existing call graph construction algorithms suffer from incompleteness
%, imprecision, or both.  
To alleviate this problem, we proposed a reflection-aware call graph
construction algorithm in Section~\ref{sec:cg}.

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that actually correspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and redundant reports, and provide
annotation support for users to specify call relationships for
 native methods.

\begin{figure}[t]
\textbf{\underline{TODO: reformat it using mathit}}\\
\textbf{Input}: a Java program $P$\\
\textbf{Output}: a set of potential $errors$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE errors $\leftarrow$ $\emptyset$
\STATE cg $\leftarrow$ $constructCallGraph$($P$)
\STATE entryNodes $\leftarrow$ $getEntryNodes$(cg)
\STATE uiAccessNodes $\leftarrow$ $getUIAccessNodes$(cg)
\STATE safeNodes $\leftarrow$ $getSafeNodes$(cg)
\FOR{each entryNode in entryNodes}
\STATE threadStartNodes $\leftarrow$ $getReachableStartsByBFS$(cg)
\FOR{each threadStart in threadStartNodes}
\STATE queue $\leftarrow$ an empty queue
\STATE queue.enqueue(threadStart.$getAllSuccNodes$())
\WHILE{queue.isNotEmpty()}
\STATE nextNode $\leftarrow$ queue.dequeue()
\FOR{each succNode in cg.$getSuccNodes$(nextNode)}
\IF{succNode $\in$ uiAccessNodes}
\STATE errors $\leftarrow$ errors $\cup$ $createErrorReport$(succNode)
\ELSIF{succNode $\in$ safeNodes}
\STATE continue
\ELSE
\STATE queue.enqueue(succNode)
\ENDIF 
\ENDFOR
\ENDWHILE
\ENDFOR
\ENDFOR
\RETURN $errors$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{Algorithm for detecting invalid thread access errors in multithreaded GUI programs. 
Any call graph construction algorithm can be used (line 2).
Three utility methods in lines 3 -- 5 are specific for each GUI framework
 as described in Section~\ref{sec:platforms}.
} \label{fig:detectalgorithm}
\end{figure}

@@

\begin{figure}[t]
\textbf{\underline{TODO: reformat it using mathit}}\\
\textbf{Input}: a Java program $P$\\
\textbf{Output}: a call graph $cg$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\FOR{each $statement$ in $P$}
\IF{$statement$ is a reflection call}
\STATE objectSet $\leftarrow$ getAllObjectsToCreate($statement$)
\STATE $newInstr$ $\leftarrow$ createObjectCreationStatement(objectSet)
\STATE replace $statement$ with $newInstr$
\ENDIF
\ENDFOR
\STATE $cg$ $\leftarrow$ $constructCallGraph$($P$, callSiteMap)
\RETURN $cg$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{A reflection-aware call graph construction algorithm. Lines 1
- 7 shows a simple program transformation to replace reflection calls
with object creations, and line 8 constructs the call graph on
the transformed program using the existing call graph algorithm. 
} 
\label{fig:cgalgorithm}
\end{figure}

Unfortunately, GUI applications often use reflection extensively and make
existing call graph construction algorithms fail to build a sufficiently complete
call graph. The example code in Figure~\ref{fig:sampleandroid} from an Android application
 illustrates the limitations.
In Figure~\ref{fig:sampleandroid}, line 6 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. When this button
is clicked, its event handling code (line 9 -- 11) updates the text.
%d caches
%the current view to a list declared at line 2.

@@

To address this limitation, we present an algorithm to construct
a reflection-aware call graph based on simple program transformation.
The basic idea to replace reflection calls with explicit object
creation instructions, pretending that the corresponding concrete object
has been created.  The algorithm is shown in Figure~\ref{fig:cgalgorithm}.
It consists of two steps. The first step is a simple program
transformation to replace reflection calls with object creations, and
the second step is to use standard call graph construction algorithms
to construct the graph on the transformed program. Lines 1 - 7 show
the first step of program transformation. The algorithm iterates through every statement
in an analyzed program. When it sees a reflection call statement (e.g.,
the \CodeIn{findViewById} method), the algorithm analyzes the
program to determine a set of possible classes that may be created.
After that, the algorithm creates a new instruction which non-deterministically
returns a new object from the class set. Line 8 is the second step
to build a call graph on the transformed program.  In this step,
any existing call graph construction algorithms can be used.
%construction algorithm is independent of the underlying call graph algorithm;
%any existing algorithm can be used to construct the resulting graph at line 8.

For the example in Figure~\ref{fig:sampleandroid}, the algorithm
first parses the XML configuration file to determine that the \CodeIn{button\_id}
value is mapped to a \CodeIn{Button} instance declaration. Then it
replaces the reflection call: \CodeIn{findViewById(R.id.button\_id)} 
with an explicit object creation call: \CodeIn{new Button()}.
After that, the algorithm employs existing call graph construction
algorithms to analyze the transformed program, and permits them
to add the call edge \CodeIn{setText} to the resulting graph.

Using this reflection-aware call graph construction algorithm is
critical to detect bugs on realistic GUI applications.
As demonstrated in our experiments, errors in some Android applications
can only be detected using this algorithm.

@@

A GUI application may also use native methods to interact with the underlying
operating system or platforms. However, native methods are often
beyond the ability of a static analysis but should be considered to make
the call graph more complete. To do so, we provide an annotation \CodeIn{@CallByNativeMethods}
for users to specify which native methods may call the current method. For example,
the following code snippet indicates that native methods \CodeIn{native1()} and \CodeIn{native2()}
will call method \CodeIn{javaMethod()}.

%\noindent 
\CodeIn{@CallByNativeMethods(callers=\{"native1", "native2"\})}

@@

Our static analysis  takes the call relationship specified by this
annotation into consideration when traversing the call graph. In our experiment,
we found such an annotation was useful: one real-world bug can only be reported
with the user-provided annotations.

@@

A static analysis can exhaustively check possible error path; but may report
paths that do not actually exist (i.e., false positives) or multiple paths
 that have the same error cause (i.e., redundant warnings). We devise
5 heuristic filters below to remove likely false positives and redundant warnings.
The first two are for eliminating false positives, while the
other three are for reducing redundant warnings.

@@

%\item
\textbf{1. Filter Reports Containing Library Calls}. A
method call chain involving normal library calls
like \CodeIn{Object.toString} or \CodeIn{Runtime.shutDown} are unlikely to
be buggy. In addition, a method call chain that \textit{only}
consists of library calls after the entry method is also unlikely
to reveal a valid bug, since it has no UI interaction.
We encode a list of common library calls, and remove
a reported warning if it contains one of them.
%Method call chain involves typical
%system methods like \CodeIn{toString} are unlikely to reveal a real bug.

%\item
\textbf{2. Filter Reports With User-Annotated Methods}. Besides general
library calls, users are also permitted
to explicitly annotate methods that will never trigger an error.
For example, the Android GUI framework provides a utility method 
\CodeIn{runOnUIThread} to check whether the current thread is the UI thread
before executing the code. If the current thread is not the UI thread,
this method will be executed on the UI thread via message passing.
Thus, a reported method call chain containing an annotated 
can be safely removed.
%method will never cause an
%invalid thread error and should be removed.

%\item
\textbf{3. Filter Lexically Redundant Reports}. One reported method call
chain can lexically subsume another one. For example, suppose that two
reported method call chains: \CodeIn{a()} $\rightarrow$ \CodeIn{b()}
$\rightarrow$ \CodeIn{c()} and 
\CodeIn{d()} $\rightarrow$ \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()}
both lead to a potential error, since $\CodeIn{d()}$ and \CodeIn{a()}
are two separated entry methods. The second call chain should
be removed and would not reduce the error detection capability, since
the first chain reveals the same error and is shorter 
for programmers to interpret.

@@

In fact, these two chains may have the same error root: knowning that
method \CodeIn{m()} is called by a non-UI thread is sufficient to understand
the error.
This filter compares two reported chains that have the same
$k$ ($k$ is user-settable with a default value 5) head nodes, and removes the longer one.

%\item
\textbf{5. Filter Report With the Same $k$ Tail Methods}. A method
can have multiple callers, so that method call chains with the same tail are likely
to just represent different ways to trigger the same error. This filter compares 
two reported method call chains that share the same 
$k$ ($k$ is user-settable with a default value 5) tail nodes, and removes the longer one.

@@

We instantiate our error detection techniques for four popular GUI frameworks,
namely SWT, Eclipse plugin environment, Swing, and Android.
The major framework-specific parts, corresponding to
lines 3 -- 5 in Figure~\ref{fig:detectalgorithm}, are to identify
\textbf{call graph entry nodes}), \textbf{UI-accessing nodes},
and \textbf{safe methods} for each framework.

@@

%SWT is an open source widget toolkit for Java designed to provide efficient,
%portable access to the GUI facilities of the operating systems
%on which it is implemented.
Like a normal Java program, a SWT desktop application
has a single main method as the entry point. By default,
the main method is executed in the UI thread after the GUI
is initialized. We instantiate our technique for SWT applications as follows:

@@

\item \textbf{Call graph entry nodes: } the main method

\item \textbf{UI-accessing nodes: } SWT provides explicit
runtime checking for illegal thread access. Any methods for manipulating
GUI object must call methods \CodeIn{Display.checkWidget}
or \CodeIn{Display.checkDevice} to check whether the current
thread is the UI thread or not.
Thus, we treat nodes corresponding to the methods
calling these two methods as UI-accessing nodes.
%must be invoked
%in the UI thread and as
%UI-accessing nodes.
 %methods to check whether the current thread is the UI thread or not. Thus,
%we treat all such methods that must be accessed in UI thread as the UI-accessing nodes.
% any
%methods that calls the checking method must be invoked in the UI thread.

\item \textbf{Safe methods: } SWT provides two helper methods (\CodeIn{Display.\\asyncExec}
and \CodeIn{Display.syncExec}) to execute code (a)synchronizedly on the UI thread.
We treat these two methods as safe methods, since invoking
them from any thread 
%are generally considered to be safe and
would not cause an invalid thread access error.

@@

%An eclipse plugin is an open program. 
An Eclipse plugin is built on top of the SWT framework.
However, unlike a SWT desktop application, an Eclipse plugin does not
have a main method as a single program entry point. Instead, it extends
extension points provided by the Eclispe framework to implement
functionalities. When a plugin starts, Eclipse
calls back the overridden extension points to handle the
upcoming events. We customize our technique for Eclipse plugin as follows:

@@

\item \textbf{Call graph entry nodes: } all overridden SWT
GUI event handling methods are called from the UI thread.
We treat these callback methods as the entry nodes.

%due to the complexity of the eclipse
%architecture,
%there is no rigorous way in eclipse to identify a UI thread. Here, we use a heuristic to determine call graph
%entry nodes from the UI thread. As a naming convention, source code in a plugin's \CodeIn{ui}
%package deals with the creation and maintainance of a plugin's GUI, and is thus
%invoked by the UI thread. We conservatively treat all public methods in the \CodeIn{ui}
%as call graph entry nodes.

%entry nodesall overriden methods of the classes inside
% a plugin's \CodeIn{ui} package.  Such methods are normally called by the eclipse framework
%by the UI thread. XXX This rule, though quite , works quite well in practice.

@@

%The collections classes from the original JDK (\CodeIn{Vector}
%and \CodeIn{Hashtable}) and the AWT library are designed as
%thread-safe. That is, two separate threads of execution can
%access the UI element at the same time without the developer having
%to worry about the threads interfering with one another. However, the
%safety comes at a cost. Because there is a great deal of overhead
%necessary to build thread-safe artifacts, the tend to be much slower
%than nonthread-safe alternatives.

%The design decisions for Swing changed. This does not mean the controls
%can never be accessed from multiple threads, but the developer is now
%responsible for adding code to ensure that no ill effects occur.

A Swing application has a single main method, but contains three kinds of
threads: \textit{initial thread} that executes initial application code from the main method,
the \textit{UI thread}, where all GUI manipulation code is executed,
and the \textit{worker thread} where time-consuming background tasks are executed.
After a Swing program starts, its initial thread exits and the UI thread takes charge
of the application and starts to execute all event-handling code or spawns new worker threads. 
We customized our technique for Swing as follows:
%XXXX Therefore, we can not directly
%use the main method as the call graph entry point, since it is not invoked
%by the UI thread. We customize our technique for Swing as follows:

@@

\item \textbf{Call graph entry nodes: } all overridden callback methods in 
Swing's UI listener classes are called from the UI thread. We treat those callback
methods as the entry nodes.

%Specifically, for each class, we check whether
%it is a Swing listener class. If so, we extract all its overridden callback methods as the call graph entry points.
 %Such methods are normally called by the eclipse framework
%by the UI thread. XXX This rule, though quite heuristical, works quite well in practice.

\item \textbf{UI-accessing nodes:} %as stated in Swing's documentation,
methods in Swing GUI classes are not thread safe with three exceptions (\CodeIn{repaint()},
\CodeIn{revalidate()}, and \CodeIn{invalidate()}). Thus, we
treat nodes corresponding to those thread unsafe methods %in each GUI class
as UI-accessing nodes.

\item \textbf{Safe methods: }  Swing provides two helper methods
(\CodeIn{SwingUti-\\lities.invokeLater} and \CodeIn{SwingUtilities.invokeAndWait}) to execute code in the UI thread.
We treat these two methods as safe methods.

@@

%A key tasks in instantiating
%our technique to the Swing framework is to identify its event dispatch thread.
%We use

@@

In our experience, instantiating our general technique to  a specific framework
requires moderate human effort. We wrote around 500 lines of Java code in total to achieve
the above four instantiations.

>>>>

We first give a high-level formulation of the problem, then present
the error detection algorithm. Finally, we show how to instantiate
the error algorithm for different GUI frameworks.

\subsection{Problem Formulation}

This section formulates the problem. We define the
notations of \textit{UI thread}, \textit{non-UI threads},
\textit{UI-accessing methods},
 \textit{safe methods}, and \textit{invalid thread access error}, and
two assumptions we make with regard to error detection.

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specifically, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretically ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative, or sound, static call graph is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
In an object-oriented programming language like
Java, a method is the basic logic unit in a program. Reporting
potential errors at the method level can be sufficient in most cases.
Thus, we believe using the call graph is well suited for this problem while
compared with other program representations like control flow graph or
program dependence graph.

Our algorithm in Figure~\ref{fig:detectalgorithm} %works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes, UI-accessing nodes, and safe nodes
for it on line 3, 4, and 5, respectively. Lines 3--5 are GUI framework-specific.
As an example, for a SWT desktop application, the entry nodes include
the single main method, UI-accessing nodes include all methods that
call method \CodeIn{Display.checkWidget} or \CodeIn{Display.checkDevice},
and safe methods include two SWT helper methods \CodeIn{Display.asyncExec}  and
\CodeIn{Display.syncExec} for message passing. The detailed instantiation
for each supported framework is explained in Section~\ref{sec:platforms}.

For each entry node, the algorithm uses standard Breadth First Search (BFS)
to find all reachable \CodeIn{Thread.start()} nodes (line 7).
Each \CodeIn{Thread.start()} node in a call graph indicates that a new,
non-UI thread is spawned.  After that, from each \CodeIn{Thread.start()}
node, the algorithm uses BFS again to find reachable UI-accessing methods (lines 9 -- 28).
Specifically, it uses a queue to keep track of nodes to be visited (line 9),
and also stores all visited nodes to avoid infinite loop (line 10).
The algorithm checks every successive node to be visited (line 18);
it reports an error if that node is a UI-accessing method (lines 20, 21), or skips
it if it is a safe method (lines 22, 23), or simply keeps traversing otherwise (line 25).

Note that when the algorithm determines that a UI-accessing method may be invoked
by a non-UI thread, it computes full method call chain from the entry node to
the UI-accessing node as the error report.
The method \CodeIn{createErrorReport} on line 20 does that. 
As an example, the method call chain for the error in in Figure~\ref{fig:androiderror}
is shown in Figure~\ref{fig:report}.

Our algorithm in Figure~\ref{fig:detectalgorithm} uses BFS to search
for potential error-revealing paths, since BFS permits to produce
small error reports (i.e., shorter method call chains). However,
other graph search strategies such as Depath-First Search (DFS) or
exhaustive path search can also be employed. In our evaluation (Section~\ref{sec:search}),
we empirically compared different graph search strategies, and demonstrate that
using BFS allows the algorithm to find more errors than DFS and
exhaustive path search.

\begin{figure}[t]
\textbf{Input}: a Java program $\mathit{P}$\\
\textbf{Output}: a set of potential invalid thread access $\mathit{errors}$\\
%\textbf{Auxiliary Methods:}\\
%getSuccNodes($\mathit{g}$, $\mathit{n}$): return node $\mathit{n}$'s successive
%nodes in the graph $\mathit{g}$\\
\vspace{-4mm}
\begin{algorithmic}[1]
\STATE $\mathit{errors}$ $\leftarrow$ $\emptyset$
\STATE $\mathit{cg}$ $\leftarrow$ constructCallGraph($\mathit{P}$)
\STATE $\mathit{entryNodes}$ $\leftarrow$ getEntryNodes($\mathit{cg}$)
\STATE $\mathit{uiAccessNodes}$ $\leftarrow$ getUIAccessNodes($\mathit{cg}$)
\STATE $\mathit{safeNodes}$ $\leftarrow$ getSafeNodes($\mathit{cg}$)
\FOR{each $\mathit{entryNode}$ in $\mathit{entryNodes}$}
\STATE $\mathit{threadStartNodes}$ $\leftarrow$ getReachableStartsByBFS($\mathit{cg}$)
\FOR{each $\mathit{threadStart}$ in $\mathit{threadStartNodes}$}
\STATE $\mathit{queue}$ $\leftarrow$ an empty queue\\
\COMMENT{keep all visited nodes to avoid infinite loop}
\STATE $\mathit{visited}$ $\leftarrow$ $\emptyset$
\STATE $\mathit{queue}$.enqueue(getSuccNodes($\mathit{cg}$, $\mathit{threadStart}$))
\WHILE{$\mathit{queue}$.isNotEmpty()}
\STATE $\mathit{node}$ $\leftarrow$ $\mathit{queue}$.dequeue()
\IF{$\mathit{node}$ $\in$ $\mathit{visited}$}
\STATE continue
\ENDIF
\STATE $\mathit{visited}$ $\leftarrow$ $\mathit{visited}$ $\cup$ $\mathit{node}$
\FOR{each $\mathit{succNode}$ in getSuccNodes($\mathit{cg}$, $\mathit{node}$)}
\IF{$\mathit{succNode}$ $\in$ $\mathit{uiAccessNodes}$}
\STATE $\mathit{newError}$ $\leftarrow$ createErrorReport($\mathit{entryNode, succNode}$)
\STATE $\mathit{errors}$ $\leftarrow$ $\mathit{errors}$ $\cup$ $\mathit{newError}$
\ELSIF{$\mathit{succNode}$ $\in$ $\mathit{safeNodes}$}
\STATE continue
\ELSE
\STATE $\mathit{queue}$.enqueue($\mathit{succNode}$)
\ENDIF 
\ENDFOR
\ENDWHILE
\ENDFOR
\ENDFOR
\RETURN $errors$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{Algorithm for detecting invalid thread access errors in multithreaded GUI programs. 
Any call graph construction algorithm can be used (line 2). The algorithm
is parameterized by the three methods in lines 3--5 which are specific to each GUI framework
 as described in Section~\ref{sec:platforms}.
} \label{fig:detectalgorithm}
\end{figure}

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph in the presence of reflection is non-trivial. 
In particular, GUI applications built on some popular frameworks like Android
often use configuration files and reflection to specify GUI layout.
To alleviate this problem, we next proposed a reflection-aware call graph
construction algorithm in Section~\ref{sec:cg}.

@@

\begin{figure}[t]
\textbf{Input}: a Java program $\mathit{P}$\\
\textbf{Output}: a call graph $\mathit{cg}$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\FOR{each $\mathit{statement}$ in $\mathit{P}$}
\IF{isReflectionCall($statement$)}
\STATE $\mathit{objectSet}$ $\leftarrow$ getAllObjectsThatMaybeCreated($\mathit{statement}$)
\STATE $\mathit{newStmt}$ $\leftarrow$ createObjectCreationStatement($\mathit{objectSet}$)
\STATE replace $\mathit{statement}$ with $\mathit{newStmt}$
\ENDIF
\ENDFOR
\STATE $\mathit{cg}$ $\leftarrow$ constructCallGraph($\mathit{P}$)
\RETURN $\mathit{cg}$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{A reflection-aware call graph construction algorithm. Lines
1--7 show a simple program transformation to replace reflection calls
with object creations, and line 8 builds the call graph using
the existing call graph construction algorithms. This algorithm
is parameterized by three methods on line 2, 3, and 8. How to
instantiate it for Android applications is presented in Section~\ref{sec:cg}.
%This algorithm is parameterized by the call graph construction
%algorithm on line 8, in which any existing algorithm can be used.
} 
\label{fig:cgalgorithm}
\end{figure}

Unfortunately, GUI applications often use reflection extensively and make
existing call graph construction algorithms such as RTA and k-CFA
fail to build a sufficiently complete call graph.
The example code in Figure~\ref{fig:sampleandroid} from an Android application
 illustrates the limitations.
In Figure~\ref{fig:sampleandroid}, line 6 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the associated XML file. When this button
is clicked, its event handling code (lines 9--11) updates the text.
%d caches
%the current view to a list declared at line 2.

@@

To address this limitation, we present an algorithm to construct
a reflection-aware call graph based on simple program transformation.
The basic idea to replace reflection calls with explicit object
creation statements, pretending that the corresponding concrete object
has been created.  The algorithm is shown in Figure~\ref{fig:cgalgorithm}.
It consists of two steps. The first step is a simple program
transformation to replace reflection calls with object creation statements, and
the second step is to use a standard call graph construction algorithm
to construct the graph on the transformed program. The reflection call
represents a method invocation that relies Java reflection to create
objects, such as the \CodeIn{findViewById} method in Android applications.
In Figure~\ref{fig:cgalgorithm}, lines 1--7 show
the first step of program transformation. The algorithm iterates through every statement
in an analyzed program. When it sees a reflection call, the algorithm analyzes the
program to determine a set of possible classes that might be created.
After that, the algorithm creates a new statement which non-deterministically
returns a new object from the object set. Line 8 is the second step
to build a call graph on the transformed program.  In this step,
any existing call graph construction algorithms can be used.

The algorithm in Figure~\ref{fig:cgalgorithm} is parameterized in three places:
line 2, line 3, and line 8. We instantiated it for Android applications as
follows. On line 2, the predicate $isReflectionCall$ returns true
if the $statement$ is a \CodeIn{findViewById(id)} method call. On line 3, method
\CodeIn{getAllObjectsThatMaybeCreated} parses the associated XML configuration
file in an Android application to extract the class declaration
 correspondinig to the \CodeIn{id} value. If the \CodeIn{id}
value is dynamically generated, the method will conservatively
return all subclasses of the declared type. On line 8, any existing
call graph construction algorithm can be used to build the call graph
on top of the transformed program.

For the example in Figure~\ref{fig:sampleandroid}, the algorithm
first parses the XML configuration file to determine that the \CodeIn{button\_id}
value is mapped to a \CodeIn{Button} instance declaration. Then it
replaces the reflection call: \CodeIn{findViewById(R.id.button\_id)} 
with an explicit object creation call: \CodeIn{new Button(null)}.
After that, the algorithm employs existing call graph construction
algorithms to analyze the transformed program, and permits them
to add the call edge \CodeIn{setText} to the resulting graph.

As demonstrated in our experiments, using this reflection-aware call
graph construction algorithm is critical to detect errors
in Android applications.

@@

A GUI application may also use native methods to interact with the underlying
operating system or platforms. However, native methods are often
beyond the ability of a static analysis but should be considered to make
the call graph more complete. To do so, we provide an annotation \CodeIn{@CalledByNativeMethods}
for users to specify which native methods may call the current method. For example,
the following code snippet indicates that native methods \CodeIn{native1()} and \CodeIn{native2()}
may call method \CodeIn{javaMethod()}.

%\noindent 
\CodeIn{@CalledByNativeMethods(callers=\{"native1", "native2"\})}

@@

Our static analysis takes the call relationship specified by this
annotation into consideration when traversing the call graph. 
Adding annotations for native methods is optional and requires manual effort.
In our experiments, 1 out of 9 programs uses native methods to interact with
the underlying operating system, and we added 3 annotations for it. We manually
searched the source files, and inspected the code to determine target methods
that a native method will call, if there are any. We found this annotation,
although requires manual effort to add, was useful: one real-world bug
can only be reported with using the user-provided annotations.

@@

A static analysis can exhaustively check possible error paths; but may report
paths that do not actually exist (false positives) or multiple paths
 that have the same error cause (redundant warnings). We devise
5 heuristic filters below to remove likely false positives and redundant warnings.
The first two are for eliminating false positives, while the
other three are for reducing redundant warnings.

@@

%\item
\textbf{1. Filter Reports Containing Library Calls}. A
method call chain containing certain library calls
like \CodeIn{Runtime.}\CodeIn{shutDown} are unlikely to
be buggy. For example, the \CodeIn{Runtime.}\CodeIn{shutDown}
method is called when JVM terminates, and uses multithreading
to dispose all GUI objects after the program exits.
We encoded a list of common library calls by proposing
a basic set of library calls that seemed natural and
generally applicable, based on our programming experience.
We later added other calls we found helpful in analyzing
programs and that we believed would be generally useful.
%We encode a list of common library calls, and remove
%a reported warning if it contains one of them.
%Method call chain involves typical
%system methods like \CodeIn{toString} are unlikely to reveal a real bug.

%\item
\textbf{2. Filter Reports With User-Annotated Methods}. Besides general
library calls, users are also permitted to explicitly annotate methods
that will never trigger an error.
For example, the Android GUI framework provides a utility method 
\CodeIn{runOnUIThread} to check whether the current thread is the UI thread
before executing the code. If the current thread is not the UI thread,
this method will be executed on the UI thread via message passing.
Thus, a reported method call chain containing an annotated 
can be safely removed. We produced the list of annotated methods
during analyzing programs on different GUI frameworks; we did this
only between experiments rather than biasing experiments by tuning
our tool to specific programs.

%\item
\textbf{3. Filter Lexically Redundant Reports}. One reported method call
chain can lexically subsume another one. For example, suppose that two
reported method call chains: \CodeIn{a()} $\rightarrow$ \CodeIn{b()}
$\rightarrow$ \CodeIn{c()} and 
\CodeIn{d()} $\rightarrow$ \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()}
both lead to a potential error, since $\CodeIn{d()}$ and \CodeIn{a()}
are two distinct entry methods. The second call chain should
be removed and would not reduce the error detection capability, since
the first chain reveals the same error and is shorter 
for programmers to interpret.

@@

In fact, these two chains may have the same error root: knowning that
method \CodeIn{m()} is called by a non-UI thread is sufficient to understand
the error.
This filter compares two reported chains that have the same
$k$ ($k$ is user-settable with a default value 5 as used in our
experiments) head nodes, and removes the longer one.

%\item
\textbf{5. Filter Report With the Same $k$ Tail Methods}. A method
can have multiple callers, so that method call chains with the same tail are likely
to just represent different ways to trigger the same error. This filter compares 
two reported method call chains that share the same 
$k$ ($k$ is user-settable with a default value 5 as used in our
experiments) tail nodes, and removes the longer one.

@@

We instantiated our error detection techniques for four popular GUI frameworks,
namely SWT, Eclipse plugin environment, Swing, and Android.
The major framework-specific parts, corresponding to
lines 3 -- 5 in Figure~\ref{fig:detectalgorithm}, are identifying
\textbf{call graph entry nodes}, \textbf{UI-accessing nodes},
and \textbf{safe methods} for each framework.

@@

%SWT is an open source widget toolkit for Java designed to provide efficient,
%portable access to the GUI facilities of the operating systems
%on which it is implemented.
 We instantiate our technique for SWT applications as follows:

@@

\item \textbf{Call graph entry nodes: } the main method. Like a normal Java program,
a SWT desktop application has a single main method as the entry point. By default,
the main method is executed in the UI thread after the GUI
is initialized.

\item \textbf{UI-accessing nodes: } SWT provides explicit
runtime checking for illegal thread access. Any methods for manipulating
GUI object must call methods \CodeIn{Display.checkWidget}
or \CodeIn{Display.checkDevice} to check whether the current
thread is the UI thread or not.
Thus, we treat nodes (in both user code and framework code) corresponding to the methods
calling these two methods as UI-accessing nodes.
%must be invoked
%in the UI thread and as
%UI-accessing nodes.
 %methods to check whether the current thread is the UI thread or not. Thus,
%we treat all such methods that must be accessed in UI thread as the UI-accessing nodes.
% any
%methods that calls the checking method must be invoked in the UI thread.

\item \textbf{Safe methods: } SWT provides two helper methods (\CodeIn{Display.\\asyncExec}
and \CodeIn{Display.syncExec}) to execute code (a)synchro-
nizedly on the UI thread.
 We treat these two methods as safe methods, since invoking
them from any thread 
%are generally considered to be safe and
will not cause an invalid thread access error.

@@

We instantiated our technique for Eclipse plugin as follows:

@@

\item \textbf{Call graph entry nodes: } all overridden SWT
GUI event handling methods in user code. An Eclipse plugin is built
on top of the SWT framework. However, unlike a SWT desktop application,
an Eclipse plugin does not have a main method as a single program entry point.
Instead, it extends extension points provided by the Eclispe framework to
implement functionalities. When a plugin starts, Eclipse
calls back the overridden extension points to handle the
upcoming events. All SWT GUI event handling methods (i.e., the overriden
methods in the class that implements \CodeIn{org.eclipse.swt.internal.\\SWTEventListener}) are
called back from the UI thread, thus are used as call graph entry nodes.

@@

A Swing application has a single main method, but contains three kinds of
threads: \textit{initial thread} that executes initial application code from the main method,
the \textit{UI thread}, where all GUI manipulation code is executed,
and the \textit{worker thread} where time-consuming background tasks are executed.
After a Swing program starts, its initial thread exits and the UI thread takes charge
of the application and starts to execute all event-handling code or spawns new worker threads. 
We instantiated our technique for Swing as follows:
%XXXX Therefore, we can not directly
%use the main method as the call graph entry point, since it is not invoked
%by the UI thread. We customize our technique for Swing as follows:

@@

\item \textbf{Call graph entry nodes:} all overridden Swing GUI event handling
methods in user code. Those event handling methods are always
called back from the UI thread, thus are used as call graph entry nodes.

\item \textbf{UI-accessing nodes:} %as stated in Swing's documentation,
all methods defined in each Swing GUI class except for three thread-safe
methods: \CodeIn{repaint()}, \CodeIn{revalidate()}, and \CodeIn{invalidate()}.

%methods in Swing GUI classes are not thread safe with three exceptions (\CodeIn{repaint()},
%\CodeIn{revalidate()}, and \CodeIn{invalidate()}). Thus, we
%treat nodes corresponding to those thread unsafe methods %in each GUI class
%as UI-accessing nodes.

\item \textbf{Safe methods: }  Swing's two helper methods
(\CodeIn{SwingUtilities.} \CodeIn{invokeLater} and \CodeIn{SwingUtilities.invokeAndWait}) that execute code in the UI thread.

@@

@@

Instantiating our general technique to  a specific framework
requires moderate human effort. We wrote around 500 lines of Java code in total to achieve
the above four instantiations.
